Regional and Urban Economics

### 2023-02-19 MSB106_Assignment.qmd
.

```{r, message=FALSE}
library(rgdal)
library(dplyr)
library(RSQLite)
library(sf)
library(tidyverse)
library(readr)
library(ggplot2)
library(kableExtra)
library(huxtable)
```

```{r, message=FALSE}
NOR_CBD <- read_csv("NOR_CBD.csv")
Dist_CBD_Dentist <- read_csv("Dist_CBD_Dentist.csv")
Zonal_CBD <- read_csv("Zonal_CBD.csv")
Dist_Mal_Dentist <- read.csv("Dist_Mal_Dentist.csv")
NOR_OSM_SHOP_MAL <- read.csv("NOR_OSM_SHOP_MAL.csv") |>
  rename("HubName" = fid)
Zonal_Den <- read.csv("Zonal_Den.csv") |>
  select(fid, X_sum)
Dist_Comp_Den <- read.csv("Dist_Comp_Den.csv") |>
  select(fid, distance)
NOR_KOMM <- read_csv("NOR_KOMM.csv")
NOR_KOMM_2 <- read_csv("NOR_KOMM_2.csv")
```

```{r}
Commune_Data <- inner_join(NOR_KOMM, NOR_KOMM_2, by = "kommunenummer") |>
  select(kommunenummer, HubName, HubDist, Turnover_capita_retail_Omsetning) |>
  rename("knr" = kommunenummer, "DistMal" = HubDist, "Turnover_capita_retail" = Turnover_capita_retail_Omsetning)
```

```{r}
Commune_Data <- inner_join(Commune_Data, NOR_OSM_SHOP_MAL, by = "HubName") |>
  select(knr, DistMal, HubName, Turnover_capita_retail, shop, Vinmonopolet) |>
  rename("SizeMall" = shop, "Winemonopoly" = Vinmonopolet)
```

```{r}
Commune_Data$SizeMall <- as.numeric(Commune_Data$SizeMall)
Commune_Data$Winemonopoly <- as.integer(as.logical(Commune_Data$Winemonopoly))
```

```{r}
Dentist_Data <- inner_join(Dist_Mal_Dentist, Dist_CBD_Dentist, by = "fid") |>
  select(fid, Juridisk.n, Antall.ans, Sum.Drifts, Sum.salgsi, Driftsresu.y, osm_id.y, latlong.y, HubName.y, HubDist.y, HubName.x, HubDist.x) |>
  rename("DistMal" = HubDist.x, "DistCBD" = HubDist.y, "HubName" = HubName.x, "HubNameCBD" = HubName.y, "Driftsres" = Driftsresu.y, "osm_id" = osm_id.y, "latlong" = latlong.y)
```

```{r}
Dentist_Data <- inner_join(Dentist_Data, Zonal_Den, by = "fid")
```

```{r}
Dentist_Data <- inner_join(Dentist_Data, Dist_Comp_Den, by = "fid") |>
  rename("DistComp" = distance)
```

```{r}
Dentist_Data <- left_join(Dentist_Data, NOR_OSM_SHOP_MAL, by = "HubName") |>
  select(fid, Juridisk.n, Antall.ans, Driftsres, Sum.salgsi, latlong, DistCBD, DistMal, X_sum, DistComp, shop, Vinmonopolet) |>
  rename("SizeMall" = shop, "Winemonopoly" = Vinmonopolet)
```

```{r}
Dentist_Data$SizeMall <- as.numeric(Dentist_Data$SizeMall)
Dentist_Data$Winemonopoly <- as.integer(as.logical(Dentist_Data$Winemonopoly))
```

```{r}
kable(summary(Commune_Data))
```

```{r}
kable(summary(Dentist_Data[3:6]))
```

```{r}
kable(summary(Dentist_Data[7:10]))
```

```{r}
kable(summary(Dentist_Data[11:12]))
```

```{r}
Commune_Data |>
  ggplot(aes(x = SizeMall, y = Turnover_capita_retail, colour = SizeMall)) +
  geom_point(lwd = .9) +
  labs(x = "Distance from Mall", y = "Turnover capita in retail per inhabitant") +
  theme_bw()
```

```{r}
Dentist_Data |>
  ggplot(aes(x = DistCBD, y = Sum.salgsi, colour = DistCBD)) +
  geom_point(lwd = .9) +
  labs(x = "Distance from CBD", y = "Income") +
  theme_bw()
```

```{r}
Dentist_Data_LIMIT <- Dentist_Data |>
  filter(Sum.salgsi < 100000000, DistCBD < 250,
         Driftsres < 20000000, Driftsres > -5000000, DistMal < 80)
```

```{r}
Dentist_Data_LIMIT |>
  ggplot(aes(x = DistCBD, y = Sum.salgsi, colour = DistCBD)) +
  geom_point(lwd = .9) + labs(x = "Distance from CBD", y = "Income") +
  theme_bw()
```

```{r}
Dentist_Data |>
  ggplot(aes(x = DistMal, y = Driftsres, colour = DistMal)) +
  geom_point(lwd = .9) +
  labs(x = "Distance from Mall", y = "EBIT") +
  theme_bw()
```

```{r}
Dentist_Data_LIMIT |>
  ggplot(aes(x = DistMal, y = Driftsres, colour = DistMal)) +
  geom_point(lwd = .9) +
  labs(x = "Distance from Mall", y = "EBIT") +
  theme_bw()
```

```{r}
Dentist_Data_LIMIT |>
  ggplot(aes(x = X_sum, y = Sum.salgsi, colour = Sum.salgsi)) +
  geom_point(lwd = .9) +
  labs(x = "Pop density", y = "Income") +
  theme_bw()
```

```{r}
lm1 <- lm(Sum.salgsi ~ DistComp + DistMal + SizeMall + (DistMal*SizeMall) + DistCBD + X_sum + (DistCBD*X_sum), data = Dentist_Data_LIMIT)
```

```{r}
huxreg(list("Dentist Revenue" = lm1), statistics = c(N = "nobs", R2 = "r.squared"), const. = "(Intercept)", note = "Note: {stars} T statistics in brackets.")
```

```{r}
Den_W_Wine <- Dentist_Data_LIMIT |>
  filter(Winemonopoly == 1)
```

```{r}
Den_WO_Wine <- Dentist_Data_LIMIT |>
  filter(Winemonopoly == 0)
```

```{r}
lm2 <- lm(Turnover_capita_retail ~ DistMal + SizeMall + (DistMal*SizeMall) + Winemonopoly, data = Commune_Data)
```

```{r}
huxreg(list("Turnover per capita in retail" = lm2), statistics = c(N = "nobs", R2 = "r.squared"), const. = "(Intercept)", note = "Note: {stars} T statistics in brackets.")
```

>>>>

```{r, message=FALSE, echo=FALSE}
library(rgdal)
library(dplyr)
library(RSQLite)
library(sf)
library(tidyverse)
library(readr)
library(ggplot2)
library(kableExtra)
library(huxtable)
```

```{r, message=FALSE, echo=FALSE}
NOR_CBD <- read_csv("NOR_CBD.csv")
Dist_CBD_Dentist <- read_csv("Dist_CBD_Dentist.csv")
Zonal_CBD <- read_csv("Zonal_CBD.csv")
Dist_Mal_Dentist <- read.csv("Dist_Mal_Dentist.csv")
NOR_OSM_SHOP_MAL <- read.csv("NOR_OSM_SHOP_MAL.csv") |>
  rename("HubName" = fid)
Zonal_Den <- read.csv("Zonal_Den.csv") |>
  select(fid, X_sum)
Dist_Comp_Den <- read.csv("Dist_Comp_Den.csv") |>
  select(fid, distance)
NOR_KOMM <- read_csv("NOR_KOMM.csv")
NOR_KOMM_2 <- read_csv("NOR_KOMM_2.csv")
```

```{r, echo=FALSE}
Commune_Data <- inner_join(NOR_KOMM, NOR_KOMM_2, by = "kommunenummer") |>
  select(kommunenummer, HubName, HubDist, Turnover_capita_retail_Omsetning) |>
  rename("knr" = kommunenummer, "DistMal" = HubDist, "Turnover_capita_retail" = Turnover_capita_retail_Omsetning)
```

```{r, echo=FALSE}
Commune_Data <- inner_join(Commune_Data, NOR_OSM_SHOP_MAL, by = "HubName") |>
  select(knr, DistMal, HubName, Turnover_capita_retail, shop, Vinmonopolet) |>
  rename("SizeMall" = shop, "Winemonopoly" = Vinmonopolet)
```

```{r, echo=FALSE, warning=FALSE}
Commune_Data$SizeMall <- as.numeric(Commune_Data$SizeMall)
Commune_Data$Winemonopoly <- as.integer(as.logical(Commune_Data$Winemonopoly))
```

```{r, echo=FALSE}
Dentist_Data <- inner_join(Dist_Mal_Dentist, Dist_CBD_Dentist, by = "fid") |>
  select(fid, Juridisk.n, Antall.ans, Sum.Drifts, Sum.salgsi, Driftsresu.y, osm_id.y, latlong.y, HubName.y, HubDist.y, HubName.x, HubDist.x) |>
  rename("DistMal" = HubDist.x, "DistCBD" = HubDist.y, "HubName" = HubName.x, "HubNameCBD" = HubName.y, "Driftsres" = Driftsresu.y, "osm_id" = osm_id.y, "latlong" = latlong.y)
```

```{r, echo=FALSE}
Dentist_Data <- inner_join(Dentist_Data, Zonal_Den, by = "fid")
```

```{r, echo=FALSE}
Dentist_Data <- inner_join(Dentist_Data, Dist_Comp_Den, by = "fid") |>
  rename("DistComp" = distance)
```

```{r, echo=FALSE}
Dentist_Data <- left_join(Dentist_Data, NOR_OSM_SHOP_MAL, by = "HubName") |>
  select(fid, Juridisk.n, Antall.ans, Driftsres, Sum.salgsi, latlong, DistCBD, DistMal, X_sum, DistComp, shop, Vinmonopolet) |>
  rename("SizeMall" = shop, "Winemonopoly" = Vinmonopolet)
```

```{r, echo=FALSE, warning=FALSE}
Dentist_Data$SizeMall <- as.numeric(Dentist_Data$SizeMall)
Dentist_Data$Winemonopoly <- as.integer(as.logical(Dentist_Data$Winemonopoly))
```

```{r, echo=FALSE}
kable(summary(Commune_Data))
```

```{r, echo=FALSE}
kable(summary(Dentist_Data[3:6]))
```

```{r, echo=FALSE}
kable(summary(Dentist_Data[7:10]))
```

```{r, echo=FALSE}
kable(summary(Dentist_Data[11:12]))
```

```{r, echo=FALSE}
#| label: fig-1
#| tbl-cap: "Capita in retail compared to mall distance"
Commune_Data |>
  ggplot(aes(x = SizeMall, y = Turnover_capita_retail, colour = SizeMall)) +
  geom_point(lwd = .9) +
  labs(x = "Distance from Mall", y = "Turnover capita in retail per inhabitant") +
  theme_bw()
```

```{r,echo=FALSE}
#| label: fig-2
#| tbl-cap: "Dentist income compared to CBD distance"
Dentist_Data |>
  ggplot(aes(x = DistCBD, y = Sum.salgsi, colour = DistCBD)) +
  geom_point(lwd = .9) +
  labs(x = "Distance from CBD", y = "Income") +
  theme_bw()
```

```{r,echo=FALSE}
Dentist_Data_LIMIT <- Dentist_Data |>
  filter(Sum.salgsi < 100000000, DistCBD < 250,
         Driftsres < 20000000, Driftsres > -5000000, DistMal < 80)
```

```{r, echo=FALSE}
Dentist_Data_LIMIT |>
  ggplot(aes(x = DistCBD, y = Sum.salgsi, colour = DistCBD)) +
  geom_point(lwd = .9) + labs(x = "Distance from CBD", y = "Income") +
  theme_bw()
```

```{r, echo=FALSE}
Dentist_Data |>
  ggplot(aes(x = DistMal, y = Driftsres, colour = DistMal)) +
  geom_point(lwd = .9) +
  labs(x = "Distance from Mall", y = "EBIT") +
  theme_bw()
```

```{r, echo=FALSE}
Dentist_Data_LIMIT |>
  ggplot(aes(x = DistMal, y = Driftsres, colour = DistMal)) +
  geom_point(lwd = .9) +
  labs(x = "Distance from Mall", y = "EBIT") +
  theme_bw()
```

```{r, echo=FALSE}
Dentist_Data_LIMIT |>
  ggplot(aes(x = X_sum, y = Sum.salgsi, colour = Sum.salgsi)) +
  geom_point(lwd = .9) +
  labs(x = "Pop density", y = "Income") +
  theme_bw()
```

```{r, echo=FALSE}
lm1 <- lm(Sum.salgsi ~ DistComp + DistMal + SizeMall + (DistMal*SizeMall) + DistCBD + X_sum + (DistCBD*X_sum), data = Dentist_Data_LIMIT)
```

```{r, echo=FALSE}
huxreg(list("Dentist Revenue" = lm1), statistics = c(N = "nobs", R2 = "r.squared"), const. = "(Intercept)", note = "Note: {stars} T statistics in brackets.")
```

```{r, echo=FALSE}
Den_W_Wine <- Dentist_Data_LIMIT |>
  filter(Winemonopoly == 1)
```

```{r, echo=FALSE}
Den_WO_Wine <- Dentist_Data_LIMIT |>
  filter(Winemonopoly == 0)
```

```{r, echo=FALSE}
lm2 <- lm(Turnover_capita_retail ~ DistMal + SizeMall + (DistMal*SizeMall) + Winemonopoly, data = Commune_Data)
```

```{r, echo=FALSE}
huxreg(list("Turnover per capita in retail" = lm2), statistics = c(N = "nobs", R2 = "r.squared"), const. = "(Intercept)", note = "Note: {stars} T statistics in brackets.")
```

### 2023-02-19 Retail Agglomeration.qmd
.

We furthermore have used plots to look at the relationship between the income of dentist\'s businesses and the distance from CBDs.
In the first scatterplot we see some extreme figures of 1 billion+ income which we choose to ignore and deem as outliers, as we notice that these figures are accounting figures for an entire group of companies under the same legal name.
By ignoring these outliers we are then left with the scatterplot in @fig-3. This scatterplot shows a tendency of a negative relationship between the variables, which might suggest that dentists might earn more closer to CBDs.
The last scatterplots we have chosen to calculate is concerning the relationship of dentists EBIT which is earnings before interest and taxes, compared to their distances to the nearest shopping mall.
Just as the previous scatterplot, we also see here a tendency of a negative relationship between the variables.
A decent number of dentists run at a loss, having negative earnings before interest and taxes.

>>>>

We furthermore have used plots to look at the relationship between the income of dentist\'s businesses and the distance from CBDs.
The scatterplot in @fig-2 have some extreme figures of 1 billion+ income which we choose to ignore and deem as outliers, as we notice that these observations are accounting figures for an entire group of companies under the same legal name, and not individual companies.
By ignoring these outliers we are then left with the scatterplot in @fig-3. This plot shows a tendency of a negative relationship between the variables, which might suggest that dentists might earn more closer to CBDs.
The last scatterplots we have chosen to calculate is concerning the relationship of dentists EBIT which is earnings before interest and taxes, compared to their distances to the nearest shopping mall in @fig-4. Just as the previous scatterplot, we also see here a tendency of a negative relationship between the variables.
A decent number of dentists run at a loss, having negative earnings before interest and taxes.
